---
title: "Moonee Valley Health Profile 2020"
output:
  word_document:
    reference_docx: template.docx
---

```{r, include = FALSE}
# libraries
pacman::p_load(tidyverse, janitor, readxl, ggpubr, flextable, officer)

#knit options - turns off everything
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.width = 7)

# colours
fair_red <-  #E55048
  
colour_table <- tibble(theme = c("Fair", "Thriving", "Connected", "Green", "Beautiful"),
                       col_code = c("#E55048", "#31788F", "#6A4479", "#4EA546", "#E3A51E"))
```


# Introduction

# Death

```{r}
# aihw data
death_prem_raw <- read_excel("data_in/death_data_aihw.xlsx", sheet = "deaths_cleaner", range = "A5:E15") %>% 
  clean_names()

death_prem <- death_prem_raw %>% 
  pivot_longer(-(name:year), names_to = "gender", values_to = "prem_deaths") %>% 
  mutate(gender = str_to_title(gender)) %>% 
  mutate(name = factor(name, levels = c("Moonee Valley (C)", "Greater Melbourne")))

death_causes <- read_excel("data_in/death_data_aihw.xlsx", sheet = "deaths_cleaner", range = "A20:F40") %>% 
  clean_names()

```

## Premature deaths

Premature deaths in Moonee Valley declined between 2014 and 2018^[Australian Institute of Health and Welfare, 2020, *Mortality Over Regions and Time (MORT) books*, 7 August 2020, Australian Government, Canberra.]. There were 142.8 premature deaths per 100,000 in Moonee Valley in 2018, compared to 160.5 for Greater Melbourne.

```{r}
ggplot(death_prem, aes(year, prem_deaths, col = gender)) +
  theme_pubclean() +
  geom_point() + geom_line() + facet_wrap(~name) + ylim(0, 300) + 
  labs(title="Premature deaths age-standardised rate (per 100,000)",
       subtitle = "Premature deaths (aged under 75)",
       x = NULL, y = NULL, colour = NULL,
       caption = "Source: Australian Institure of Health and Welfare, 2020") +
  theme(legend.position = "bottom")
```
\pagebreak

# Top causes of death

The most common causes of death in 2018 were Coronary heart disease, Dementia and Cerebrovascular disease.

```{r}
# Top ten causes of death in the City of Moonee Valley, 2018
death_causes_data <- death_causes %>% 
  select(cause_of_death, deaths, per_cent_of_all_causes) %>% 
  head(10) %>% 
  mutate(per_cent_of_all_causes = round(per_cent_of_all_causes, 1)) %>% 
  mutate(deaths = round(deaths, 1)) %>% 
  rename(`Cause` = cause_of_death, Deaths = deaths, `Per cent of all causes` = per_cent_of_all_causes)

death_causes_data %>% 
  flextable()  %>% 
  autofit() %>% 
  theme_vanilla() %>% 
  footnote( i = 1, j = 1:3,
            value = as_paragraph(
              c("Source: Australian Institure of Health and Welfare, 2020")
            ),
            ref_symbols = c(""),
            part = "header")

```

# Victorian Population Health Survey 2017

The graph below compares the City of Moonee Valley to other local government areas for a number of risk factors, chronic diseases and screening tests. The colours indicate how Moonee Valley compares to Victoria. 

```{r}
# top half
vphs_snap1_raw <- read_excel("data_in/Victorian Population Health Survey (VPHS) 2017 Quick statistics for local government areas and time series data.xlsx", sheet = "LGA snapshot", range = "A4:GI52") %>% 
  clean_names() %>% 
  remove_empty() %>% 
  mutate(x2 = if_else(is.na(x2), "Victoria", x2)) %>% 
  rename(lga = x2) %>% 
  select(-local_government_area) %>% 
  select(-starts_with("x"))

# bottom half
vphs_snap2_raw <- read_excel("data_in/Victorian Population Health Survey (VPHS) 2017 Quick statistics for local government areas and time series data.xlsx", sheet = "LGA snapshot", range = "A56:GI89") %>% 
  clean_names() %>% 
  remove_empty() %>% 
  mutate(x2 = if_else(is.na(x2), "Victoria", x2)) %>% 
  rename(lga = x2) %>% 
  select(-local_government_area) %>% 
  select(-starts_with("x")) %>% 
  filter(lga != "Victoria") # only need the one from the first table
  
# join them  
vphs_snap_raw <- bind_rows(vphs_snap1_raw, vphs_snap2_raw)

# point estimates
vphs_snapshot_point_est <- vphs_snap_raw %>% 
  select(-ends_with("_2017")) %>% 
  select(-ends_with("_rse")) %>% 
  pivot_longer(-lga, names_to = "indicator", values_to = "percentage")

# indicator names
vphs_ind_names <- vphs_snapshot_point_est %>% 
  distinct(indicator) %>% 
  pull()

# Victoria point estimates
vphs_vic_point_est <- vphs_snapshot_point_est %>% 
  filter(lga == "Victoria")

# Vic lower limits
vphs_shapshot_ll <- vphs_snap_raw %>% 
  filter(lga == "Victoria") %>% 
  select(ends_with("ll_2017"))
names(vphs_shapshot_ll) <- vphs_ind_names
vphs_shapshot_lower <-  vphs_shapshot_ll %>% 
  rename_all(paste0, "_ll") %>% 
  mutate(region = "Victoria") %>% 
  pivot_longer(-region, names_to = "ind", values_to = "lower") %>% 
  pull()

# Vic upper limits
vphs_shapshot_ul <- vphs_snap_raw %>% 
  filter(lga == "Victoria") %>% 
  select(ends_with("ul_2017"))
names(vphs_shapshot_ul) <- vphs_ind_names
vphs_shapshot_upper <-  vphs_shapshot_ul %>% 
  rename_all(paste0, "_ul") %>% 
  mutate(region = "Victoria") %>% 
  pivot_longer(-region, names_to = "ind", values_to = "upper") %>% 
  pull()

# indicator groups
indicator_groups <- c(rep("Modifiable risk factors", 16), rep("Doctor diagnosed \nchronic disease", 9), rep("Screening tests \nand treatment", 6))

# join these up
vphs_vic_joined <- vphs_vic_point_est %>% 
  mutate(vic_lower = vphs_shapshot_lower) %>% 
  mutate(vic_upper = vphs_shapshot_upper) %>% 
  rename(vic_point_est = percentage) %>% 
  mutate(type = indicator_groups) %>% 
  select(-lga)

# joined to the lga data
vphs_comparison_all <- vphs_snapshot_point_est %>% 
  left_join(vphs_vic_joined, by = "indicator") %>% 
  mutate(comparison = case_when(percentage < vic_lower ~ "Significantly lower",
                                percentage > vic_upper ~ "Significantly higher",
                                TRUE ~ "Not sig. diff.")) %>% 
  mutate(indicator = str_to_sentence(indicator)) %>% 
  mutate(indicator = str_replace_all(indicator, "_", " ")) %>% 
  mutate(indicator = str_replace_all(indicator, "yearsa", "years")) %>% 
  mutate(indicator = str_replace_all(indicator, "yearsb", "years")) %>% 
  mutate(indicator = case_when(indicator == "Take away food consumed 1 day week" ~ "Take away food consumed >1 day/week",
                               indicator == "Current i e daily or occasional smoker" ~ "Current (i.e. daily or occasional smoker)",
                               indicator == "Sedentary inactive" ~ "Sedentary (inactive)",
                               indicator == "Had a faecal occult blood test 50 years of age b" ~ "Had a faecal occult blood test 50 years of age",
                               indicator == "Had a bowel exam to detect cancer last 5 years b" ~ "Had a bowel exam to detect cancer last 5 years",
                               indicator == "Increased risk of injury from a single occasion of drinking" ~ "Incr. risk of injury from a single occasion of drinking",
                               TRUE ~ indicator)) %>% 
  mutate(comparison = factor(comparison, levels = c("Significantly lower", "Not sig. diff.", "Significantly higher"))) 



#for graph
vphs_comparison <- vphs_comparison_all %>% 
  filter(lga != "Victoria") 

vphs_mv <- vphs_comparison %>% 
  filter(lga == "Moonee Valley (C)")

vphs_vic <- vphs_comparison_all %>% 
  filter(lga == "Victoria")
```

```{r, fig.width=7, fig.height=7}
ggplot(vphs_comparison, aes(fct_rev(indicator), percentage)) + 
  theme_pubclean() +
  geom_point(colour = "grey", alpha = 0.75) + coord_flip() +
  geom_point(data = vphs_mv, aes(fct_rev(indicator), percentage, col = comparison, alpha = 0.5, size = 0.05)) +
  geom_point(data = vphs_vic, aes(fct_rev(indicator), percentage, shape = lga)) +
  theme(legend.position = c(0.75, 0.89), legend.box="vertical", legend.margin=margin()) + 
  scale_colour_manual(values = c("#E55048", "#E3A51E", "#4EA546")) +
  labs(x = NULL, y = "Percentage", title = "Selected risk factors, disease and tests",
       subtitle = "City of Moonee Valley compared to other LGAs",
       caption = "Proportion of adults(18+ years). Source: Victorian Population Health Survey, 2017",
       colour = "Compared to Victoria", shape = NULL) + 
  guides(size = FALSE, alpha = FALSE) + 
  guides(col = guide_legend(nrow = 3)) +
  theme(legend.title = element_text(size=9, face="bold")) +
  theme(legend.text = element_text(size=9)) +
  facet_grid(rows = vars(type), scales = "free_y", space = "free_y")
```



# To add

* aged care https://gen-agedcaredata.gov.au/Topics/Services-and-places-in-aged-care/Explore-services-and-places-in-aged-care
* heat vulnerability index
* ahpc
* vichealth indicators
* vphs


