---
title: "Victorian Population Health Survey 2017 summary"
output:
  html_document:
    df_print: paged
---

# Introduction

This is a summary of data for the City of Moonee Valley, from the Victorian Population Health Survey 2017. The survey was conducted by the Department of Health and Human Services. More information about the survey can be found here: https://www2.health.vic.gov.au/public-health/population-health-systems/health-status-of-victorians/survey-data-and-reports/victorian-population-health-survey/victorian-population-health-survey-2017


```{r, include = FALSE}
# libraries
pacman::p_load(tidyverse, janitor, readxl, ggpubr, flextable, officer)

#knit options - turns off everything
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.width = 7)
```

```{r}
# top half
vphs_snap1_raw <- read_excel("data_in/Victorian Population Health Survey (VPHS) 2017 Quick statistics for local government areas and time series data.xlsx", sheet = "LGA snapshot", range = "A4:GI52") %>% 
  clean_names() %>% 
  remove_empty() %>% 
  mutate(x2 = if_else(is.na(x2), "Victoria", x2)) %>% 
  rename(lga = x2) %>% 
  select(-local_government_area) %>% 
  select(-starts_with("x"))

# bottom half
vphs_snap2_raw <- read_excel("data_in/Victorian Population Health Survey (VPHS) 2017 Quick statistics for local government areas and time series data.xlsx", sheet = "LGA snapshot", range = "A56:GI89") %>% 
  clean_names() %>% 
  remove_empty() %>% 
  mutate(x2 = if_else(is.na(x2), "Victoria", x2)) %>% 
  rename(lga = x2) %>% 
  select(-local_government_area) %>% 
  select(-starts_with("x")) %>% 
  filter(lga != "Victoria") # only need the one from the first table
  
# join them  
vphs_snap_raw <- bind_rows(vphs_snap1_raw, vphs_snap2_raw)

# point estimates
vphs_snapshot_point_est <- vphs_snap_raw %>% 
  select(-ends_with("_2017")) %>% 
  select(-ends_with("_rse")) %>% 
  pivot_longer(-lga, names_to = "indicator", values_to = "percentage")

# indicator names
vphs_ind_names <- vphs_snapshot_point_est %>% 
  distinct(indicator) %>% 
  pull()

# Victoria point estimates
vphs_vic_point_est <- vphs_snapshot_point_est %>% 
  filter(lga == "Victoria")

# Vic lower limits
vphs_shapshot_ll <- vphs_snap_raw %>% 
  filter(lga == "Victoria") %>% 
  select(ends_with("ll_2017"))
names(vphs_shapshot_ll) <- vphs_ind_names
vphs_shapshot_lower <-  vphs_shapshot_ll %>% 
  rename_all(paste0, "_ll") %>% 
  mutate(region = "Victoria") %>% 
  pivot_longer(-region, names_to = "ind", values_to = "lower") %>% 
  pull()

# Vic upper limits
vphs_shapshot_ul <- vphs_snap_raw %>% 
  filter(lga == "Victoria") %>% 
  select(ends_with("ul_2017"))
names(vphs_shapshot_ul) <- vphs_ind_names
vphs_shapshot_upper <-  vphs_shapshot_ul %>% 
  rename_all(paste0, "_ul") %>% 
  mutate(region = "Victoria") %>% 
  pivot_longer(-region, names_to = "ind", values_to = "upper") %>% 
  pull()

# indicator groups
indicator_groups <- c(rep("Modifiable risk factors", 16), rep("Doctor diagnosed \nchronic disease", 9), rep("Screening tests \nand treatment", 6))

# join these up
vphs_vic_joined <- vphs_vic_point_est %>% 
  mutate(vic_lower = vphs_shapshot_lower) %>% 
  mutate(vic_upper = vphs_shapshot_upper) %>% 
  rename(vic_point_est = percentage) %>% 
  mutate(type = indicator_groups) %>% 
  select(-lga)

# joined to the lga data
vphs_comparison_all <- vphs_snapshot_point_est %>% 
  left_join(vphs_vic_joined, by = "indicator") %>% 
  mutate(comparison = case_when(percentage < vic_lower ~ "Significantly lower",
                                percentage > vic_upper ~ "Significantly higher",
                                TRUE ~ "Not sig. diff.")) %>% 
  mutate(indicator = str_to_sentence(indicator)) %>% 
  mutate(indicator = str_replace_all(indicator, "_", " ")) %>% 
  mutate(indicator = str_replace_all(indicator, "yearsa", "years")) %>% 
  mutate(indicator = str_replace_all(indicator, "yearsb", "years")) %>% 
  mutate(indicator = case_when(indicator == "Take away food consumed 1 day week" ~ "Take away food consumed >1 day/week",
                               indicator == "Current i e daily or occasional smoker" ~ "Current (i.e. daily or occasional smoker)",
                               indicator == "Sedentary inactive" ~ "Sedentary (inactive)",
                               indicator == "Had a faecal occult blood test 50 years of age b" ~ "Had a faecal occult blood test 50 years of age",
                               indicator == "Had a bowel exam to detect cancer last 5 years b" ~ "Had a bowel exam to detect cancer last 5 years",
                               indicator == "Increased risk of injury from a single occasion of drinking" ~ "Incr. risk of injury from a single occasion of drinking",
                               TRUE ~ indicator)) %>% 
  mutate(comparison = factor(comparison, levels = c("Significantly lower", "Not sig. diff.", "Significantly higher")))
write_csv(vphs_comparison_all, "data_out/vphs_comparison.csv")

#for graph
vphs_comparison <- vphs_comparison_all %>% 
  filter(lga != "Victoria") 

vphs_mv <- vphs_comparison %>% 
  filter(lga == "Moonee Valley (C)")

vphs_vic <- vphs_comparison_all %>% 
  filter(lga == "Victoria")

# for table ######################################
# mv point estimates
vphs_mv_point_est <- vphs_snapshot_point_est %>% 
  filter(lga == "Moonee Valley (C)")

# mv lower limits
vphs_shapshot_mv_ll <- vphs_snap_raw %>% 
  filter(lga == "Moonee Valley (C)") %>% 
  select(ends_with("ll_2017"))
names(vphs_shapshot_mv_ll) <- vphs_ind_names
vphs_shapshot_mv_lower <-  vphs_shapshot_mv_ll %>% 
  rename_all(paste0, "_ll") %>% 
  mutate(region = "Moonee Valley (C)") %>% 
  pivot_longer(-region, names_to = "ind", values_to = "lower") %>% 
  pull()

# mv upper limits
vphs_shapshot_mv_ul <- vphs_snap_raw %>% 
  filter(lga == "Moonee Valley (C)") %>% 
  select(ends_with("ul_2017"))
names(vphs_shapshot_mv_ul) <- vphs_ind_names
vphs_shapshot_mv_upper <-  vphs_shapshot_mv_ul %>% 
  rename_all(paste0, "_ul") %>% 
  mutate(region = "mvtoria") %>% 
  pivot_longer(-region, names_to = "ind", values_to = "upper") %>% 
  pull()

# join these up
vphs_mv_joined <- vphs_mv_point_est %>% 
  mutate(lower = vphs_shapshot_mv_lower) %>% 
  mutate(upper = vphs_shapshot_mv_upper) %>% 
  rename(point_est = percentage) %>% 
  mutate(type = indicator_groups) %>% 
  select(-lga) %>% 
  mutate(indicator = str_to_sentence(indicator)) %>% 
  mutate(indicator = str_replace_all(indicator, "_", " ")) %>% 
  mutate(indicator = str_replace_all(indicator, "yearsa", "years")) %>% 
  mutate(indicator = str_replace_all(indicator, "yearsb", "years")) %>% 
  mutate(indicator = case_when(indicator == "Take away food consumed 1 day week" ~ "Take away food consumed >1 day/week",
                               indicator == "Current i e daily or occasional smoker" ~ "Current (i.e. daily or occasional smoker)",
                               indicator == "Sedentary inactive" ~ "Sedentary (inactive)",
                               indicator == "Had a faecal occult blood test 50 years of age b" ~ "Had a faecal occult blood test 50 years of age",
                               indicator == "Had a bowel exam to detect cancer last 5 years b" ~ "Had a bowel exam to detect cancer last 5 years",
                               indicator == "Increased risk of injury from a single occasion of drinking" ~ "Incr. risk of injury from a single occasion of drinking",
                               TRUE ~ indicator)) %>% 
  mutate(across(is.numeric, round, 1)) %>% 
  select(type, everything()) %>% 
  rename(Indicator = indicator, Percentage = point_est, `Lower limit` = lower, `Upper limit` = upper, Type = type) %>% 
  write_csv("data_out/vphs_comparison_mv.csv")
```


# Comparison against other local government areas

```{r, fig.width=7, fig.height=7}
ggplot(vphs_comparison, aes(fct_rev(indicator), percentage)) + 
  theme_pubclean() +
  geom_point(colour = "grey", alpha = 0.75) + coord_flip() +
  geom_point(data = vphs_mv, aes(fct_rev(indicator), percentage, col = comparison, alpha = 0.5, size = 0.05)) +
  geom_point(data = vphs_vic, aes(fct_rev(indicator), percentage, shape = lga)) +
  theme(legend.position = c(0.75, 0.89), legend.box="vertical", legend.margin=margin()) + 
  scale_colour_manual(values = c("Significantly lower" = "#E55048", "Not sig. diff." = "#E3A51E", "Significantly higher" = "#4EA546")) +
  labs(x = NULL, y = "Percentage", title = "Selected risk factors, disease and tests",
       subtitle = "City of Moonee Valley compared to other LGAs",
       caption = "Proportion of adults(18+ years). Source: Victorian Population Health Survey, 2017",
       colour = "Compared to Victoria", shape = NULL) + 
  guides(size = FALSE, alpha = FALSE) + 
  guides(col = guide_legend(nrow = 3)) +
  theme(legend.title = element_text(size=9, face="bold")) +
  theme(legend.text = element_text(size=9)) +
  facet_grid(rows = vars(type), scales = "free_y", space = "free_y")
```
\pagebreak

# Data for Moonee Valley

## Modifiable risk factors

```{r}
vphs_mv_joined %>% 
  filter(Type == "Modifiable risk factors") %>% 
  select(-Type) %>% 
  flextable() %>% 
  #merge_v(j = ~Type) %>% 
  autofit() %>% 
  theme_vanilla()
```


## Doctor diagnosed chronic disease

```{r}
vphs_mv_joined %>% 
  filter(Type == "Doctor diagnosed \nchronic disease") %>% 
  select(-Type) %>% 
  flextable() %>% 
  #merge_v(j = ~Type) %>% 
  autofit() %>% 
  theme_vanilla()
```

## Screening tests and treatment

```{r}
vphs_mv_joined %>% 
  filter(Type == "Screening tests \nand treatment") %>% 
  select(-Type) %>% 
  flextable() %>% 
  #merge_v(j = ~Type) %>% 
  autofit() %>% 
  theme_vanilla()
```
